Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-10-21 algorithm.tex
AA: exploration vs exploitation; parallelism

Observe that merging and splitting operation are dynamic processes. A portion
of the parameter space may be populated by a large number of small
regions in some eras, but can be covered by a small number of large
regions in next eras. This is a desirable property of PS, since, after a number
of eras of intense exploration of a portion of configuration space, this
exploration has turned to be enough thorough and so it may be time for
other portions to be minutely scanned.

A pseudo-code representation of PS algorithm is in Alg.\ref{alg:algo}

>>>>

\subsection{Exploration vs exploitation in PS}
Our algorithm is able to inherently balance exploration and exploitation, which are ``two cornerstones of problem solving by search''~\cite{eiben1998evolutionary}. While a detailed survey on the exploration vs exploitation trade-off can be found in~\cite{vcrepinvsek2013exploration}, here we limitedly recall that \emph{exploration} is the capability of an algorithm to consider different parts of the configuration space, not being confined only to a limited subset and being trapped in local optima. Following the terminology in \cite{weise2012evolutionary}, if the exploration is not sufficiently wide, we sat that the algorithm suffers from \emph{premature convergence}, i.e. it starts to analyze only a narrow portion of the configuration space, that may lead to ignore potential Pareto configurations that are outside this portion. When premature convergence happens, there is insufficient \emph{diversity}~\cite{weise2012evolutionary}, i.e. all the provided solutions are similar.
On the other hand, \emph{exploitation} is the capability to recognize and densely analyze the parts of configuration space that have many Pareto configurations.

As in genetic algorithms exploration and exploration are guaranteed by mutation, crossover and selection, in our algorithm these guarantees are provided for the following reasons:
	\begin{itemize}
		\item Merging and splitting operation are dynamic processes. A portion
of the parameter space may be populated by a large number of small
regions in some eras. However, after having been sufficiently explored, the innovation score of these regions may decrease and therefore that portion may be covered by a small number of large regions in next eras. This is a desirable property of PS, since, after a number of eras of intense exploration of a portion of configuration space, this exploration may turn to be enough thorough and so it may be time for other portions to be minutely scanned. Otherwise stated, PS dynamics can be summarized as follows: while eras alternate, PS zooms out from the configurations subspaces already sufficiently scanned in the past, whereas it zooms in subspaces which provide novelty (guaranteeing exploitation), meaning that they have not been properly analyzed yet (this helps exploration). 
		\item No portion of parameter space is neglected since, at each era, all the regions, even the least interesting ones, have the chance to be explored with new evaluated points (providing exploration), even though some regions are less densely explored than others.
	\end{itemize}

To summarize,thanks to the above characteristics, PS is able to recognize the regions when the best configurations are while not being trapped in only a narrow subset of configuration space.

A pseudo-code representation of PS algorithm is in Alg.\ref{alg:algo}.

### 2014-10-21 appendix.tex
AA: exploration vs exploitation; parallelism

>>>>

\subsection{Parallelism}
While a detailed analysis of the parallel capabilities of the algorithm is not in the scope of this paper, we point out that our algorithm shows parallel capabilities. Following the approach used in the ``Master-Slaves parallel genetic algorithms''~\cite{Cantu-Paz98asurvey,Borkar14parallel}, at each era a master processing node assigns to each slave processing node a set of regions. Each processing node independently performs the analysis of the assigned regions and calculates the innovation score. At the end of each era, the master collects the results from the slaves, decide performs the splitting/merging operations and the next era starts. Another possible approach is closer to the ``Island scheme'' for GA parallelization~\cite{Borkar14parallel}. The configuration space is initially divided in $n$ macro-regions, each one assigned to a different processing node. Then, each processing node performs PS algorithm on the assigned macro-regions, autonomously performing all the steps of the algorithm, including merging and splitting. Note that, while the Island scheme for GA parallelization requires a migration operator, that is not easy to properly tune, we do not have special requirements. On the other hand, one of the macro-regions may begin to provide very low innovation at a certain era. In this case, dedicating to it the same number $K$ of test function as the other macro-regions is inefficient. To solve this problems, periodically, the number $k$ of test functions assigned to each macro-region should be renegotiated, in favor of the macro-regions providing most innovation.

### 2014-10-21 experiments.tex
AA: exploration vs exploitation; parallelism

\subsection{Evaluation metrics}
In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. Both of them permit to quantify the extent of the Pareto front that is an important factor to consider when evaluating the quality of a Pareto front. Indeed, authors in \cite{zitzler_ec00} carefully observe that an exploration algorithm should preferably provide pareto fronts whose objective function range is large.
Such Pareto fronts permit to investigate in a broader way the potentials of the designed device. This is important, in general, and turns out to be essential when designing a device that may be required to work in different operational conditions, for example with stringent execution time constraints and less strict power dissipation ones or, on the contrary, with low power requirements and tolerance to high execution time.

>>>>

\subsection{Evaluation metrics}
In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. Both of them permit to quantify the extent of the Pareto front that is an important factor to consider when evaluating the quality of a Pareto front. Indeed, authors in \cite{zitzler_ec00,weise2012evolutionary} carefully observe that an exploration algorithm should preferably provide Pareto fronts whose objective function range is large. Following the terminology of~\cite{weise2012evolutionary}, a \emph{uniform convergence} should be guaranteed.
Such Pareto fronts permit to investigate in a broader way the potentials of the designed device. This is important, in general, since if only a narrow portion of the Pareto front is provided due to non-uniform convergence, the decision maker may neglect some important options~\cite{weise2012evolutionary}. Having a narrow Pareto front becomes essential, in particular, when designing a device that may be required to work in different operational conditions, for example with stringent execution time constraints and less strict power dissipation ones or, on the contrary, with low power requirements and tolerance to high execution time.
